Analysing Constraint Grammars with a SAT-solver

### 2016-03-07 paper/lrec/eval_futurework.tex
conclusions

\section{Future work and conclusions}
\label{sec:conclusion}

The evaluation indicates that the tool finds conflicts and dead rules
from actual grammars. We did not have a volunteer to test the tool in
the process of grammar writing, so we cannot conclude whether the
constructed examples are useful for getting new insights on the rules.
However, there are still a number of features to improve and add.

@@

Our solution to hardcode the tag combinations in the readings is
feasible for simple morphology, but it can cause problems with more
complex morphology.
Another benefit of this solution is the addition of lemmas and word
forms, as well as regular expressions: currently, if we want to add
one more lemma for a verb, we need to create as many new variables as
there are distinct verb forms---easily hundreds for languages with
rich morphology.
This may also be feasible to support regular expressions, in a nicer
way than just spitting out the regex. (MORE ON THAT)

\paragraph{Full expressivity of CG-3}
As for longer-term goals, we want to handle the full expressivity of CG-3,
with \textsc{map}, \textsc{add},  \textsc{addreading} and
\textsc{substitute} rules, as well as dependency structure. 
This also means finding different kinds of conflicts.
In order to implement rules that may add new readings, or new tags to
existing readings, we need to modify our approach in the SAT-encoding.
Even if the lexicon gives all readings that exist in the lexicon, the
user might give a nonexistent reading, or in the case of MAP, a
syntactic tag, which is (by definition) not in the lexicon. A more
scalable solution would be to make each tag a variable, and ask the
question ``can this reading be a noun? how about singular? how about
conditional?'' separately for each tag. Then we could lift the
restriction of tag combinations into a SAT solver: make SAT clauses
that prohibit a comparative to go with a verb, or conditional with a noun.

@@

\paragraph{Support for grammar writers}
One additional feature is to suggest reformattings for a rule. Recall
figure~\ref{fig:regroup} from the introduction; in that case, the
original rule was written by the original author, and another
grammarian thought that the latter form is nicer to read. Doing the
reverse operation could also be possible. If a rule with long
disjunctions conflicts, it may be useful to split it into smaller
conditions, and eliminate one at a time, in order to find the
reason(s) for the conflict.
As soon as the tools are mature enough, we want to
evaluate them with actual grammar writers,
in comparison with a corpus-based method or machine learning.
Finally, if the method proves feasible for CG, we want
to try applying it to other grammar formalisms.

\subsection{Conclusions}

It is nice!

>>>>

\section{Conclusions and Future Work}
\label{sec:conclusion}

We set out to design and implement an automatic analysis of constraint grammars that can find problematic rules and rule combinations, without the need for a corpus.
Our evaluation indicates that the tool indeed finds non-trivial conflicts and dead rules
from actual grammars.

We did not have a volunteer to test the tool in
the process of grammar writing, so we cannot conclude whether the
constructed examples are useful for getting new insights on the rules.

Also, there are still a number of features to improve and add.

@@

Our solution to hardcode the tag combinations in the readings is
feasible for simple morphology, but it can cause problems with more
complex morphology.
We are currently working on adding the concepts of lemmas and word
forms directly to the representation of the possible readings.
Currently, if we want to add one more lemma for a verb, we need to
create as many new variables as there are distinct verb forms---easily
hundreds for languages with rich morphology. (INARI: Do you really mean to say 'verb forms' here?)

This may also be feasible to support regular expressions, in a nicer
way than just spitting out the regex. (INARI: You said MORE ON THAT here, good because I don't understand what you want to say here :)

\paragraph{Full expressivity of CG-3}
As for longer-term goals, we want to handle the full expressivity of CG-3,
with \textsc{map}, \textsc{add},  \textsc{addreading} and
\textsc{substitute} rules, as well as dependency structure. 
This also means finding different kinds of conflicts.
In order to implement rules that may add new readings, or new tags to
existing readings, we need to modify our approach in the SAT-encoding.
Even if the lexicon gives all readings that exist in the lexicon, the
user might give a nonexistent reading, or in the case of MAP, a
syntactic tag, which is (by definition) not in the lexicon. We may need to move
to a more scalable solution.
% would be to make each tag a variable, and ask the
%question ``can this reading be a noun? how about singular? how about
%conditional?'' separately for each tag. Then we could lift the
%restriction of tag combinations into a SAT solver: make SAT clauses
%that prohibit a comparative to go with a verb, or conditional with a noun.

@@

\paragraph{Support for grammar writers}
One additional feature is to suggest reformattings for a rule. Recall
figure~\ref{fig:regroup} from the introduction; in that case, the
original rule was written by the original author, and another
grammarian thought that the latter form is nicer to read. Doing the
reverse operation could also be possible. If a rule with long
disjunctions conflicts, it may be useful to split it into smaller
conditions, and eliminate one at a time, in order to find the
reason(s) for the conflict.
Our next step is to evaluate our tools together with actual grammar writers,
in comparison with a corpus-based method or machine learning.

\paragraph{Other grammar formalisms}
Finally, we would like to investigate logic-based methods for analysing other grammar
formalisms, for example Grammatical Framework \cite{gf}. (INARI: can you add a reference here?)

%\subsection{Conclusions}

%It is nice!
